---
output: 
  md_document:
    variant: markdown_github
bibliography: paper.bib
---

# Summary

The delivery of quality health care is a constant act of balancing  demand  against  capacity,  with emerging, data intensive, artificial intelligence (AI) and machine learning (ML) approaches poised to bridge gaps in the large, resource-limited 
sector  [\@harwich2018thinking; \@wilson2019high; \@nelson2019predicting; \@yu2018artificial].  The scale of data required in such projects magnifies the importance of  existing  ethical and legal frameworks for research with human participants  [\@sales2000ethics; \@UKRI], notably around the risks posed by the processing of personally identifiable data (PID) and pseudo-PID (variables which if used together can identify an individual)  [\@ICO].  

One approach to dealing with PID concerns is to apply transformations to the data, e.g. encryption of names, or aggregation of ages, which  can limit the risk of identification  at the cost of nuance  [\@tachepun2020data].   Hence, we demonstrate an extendable package of tools for the implementation  and application of deidentification techniques to panel data sets.

# Statement of need

In order to broaden the access to sensitive data, data handlers need an auditable open process for the application of data transforms that limit the risk of personal identification (sometimes refereed to as 'data masking' transformations).  Several production scale systems exist for applying data masking transformations (e.g.  [Delphix](https://www.delphix.com/),  [K2View](https://www.k2view.com), and  [Accutive](https://accutive.com/)), but these are often expensive and cumbersome, limiting uptake in the domain of research.  The research community hence requires a simple toolbox for the application of several common 'data masking transforms' implemented  in an open source language.

This implementation of the `deident` methods is in R, chosen due to the increased adoption of open source software into the working practices of the NHS (driven by the growth on the NHS-R community and similar groups).  However, the underlying specification, and implementation of the transforms can be ported to other languages.


This package was designed considering the ICO guidelines on anonymization  [\@ICOAnonymisation] and draws on pre-existing methodologies and naming conventions  [\@NIST2015; \@integrateIO]. The package implemented de-identification via:   

*	pseudonymization – the consistent replacement of a string by a random string
* encryption - the consistent replacement of a string by an alpha-numeric hash using an encryption key and salt
*	shuffling – replacement of columns by a random sample without replacement
*	bluring – the aggregation of numeric or categorical data according to specified rules
*	perturbation – the addition of user-defined random noise to a numeric variable


Following the design principles of the existing tidyverse  [\@tidyverse] domain for ease of
adoption.  The package includes tools to create a single pipeline for the 
application of a multi-step deidentification pipeline to multiple files 
stored within the same repository, alongside the option to serialize and define 
the pipeline with yaml.  This approach allows a researcher to design and 
implement an appropriate de-identification plan and deliver it to the research 
support or business intelligence team of an organisation with limited knowledge 
of the sensitive data.  The supply of an easy method for de-identifying data
sets which requires little scripting knowledge by Trust staff may aid in 
overcoming several information governance risks that keep operational data 
siloed within health Trusts.  

# Comparison to existing R packages

Several packages have undergone development for the implementation of encryption methods to minimize identifiability within data sets (e.g. `anonymizer`  [\@Anonymizer2024], `deidentifyr`  [\@deidentifyr2024] and `digest` [\@digest2021]).  While these packages implement a variety of encryption tools, such systems are not infallible [\@wang2005break; \@szikora2022end ] especially if the data being encrypted is drawn from a known domain such as common names.  This package introduces the stateful 'pseudonymization' method by which string vectors are replaced by a randomly generated hash the first time they are observed and then preserved for re-use.  As such, breaking a single hash no longer exposes the 'key' and 'salt' the encryption relies on.  In addition, this package introduces methods for removing pseudo-identifiability (identification via the combination of features) via the other methods included.
 
# In practice

To install the current version of the `deident` package, run the command

``` {r, eval=F}
# install.packages("pak")
pak::pkg_install("Stat-Cook/deident")
```

The core functionality of the package is the `deident` function. To demonstrate functionality we use a subset of the `babynames` data set [\@babynames2021] consisting of the final two years.  

``` {r}
babynames <- babynames::babynames |> 
  dplyr::filter(year > 2015) 

str(babynames)
```


The `deident` function produces a pipeline of actions and the variables to be transformed, with each subsequent call of `deident` adding a further transformation.    The first argument of `deident` can either be a data frame or the output of a previous `deident` call.

The simplest use case is  transforming one variable via a single method:

``` {r, eval=F}
library(deident)

pipeline <- deident(babynames, "psudonymize", name)

apply_deident(babynames, pipeline)

```

The same method can be applied to multiple variables by adding the variable names:

``` {r, eval=F}
pipeline2 <- deident(babynames, "psudonymize", name, sex)

apply_deident(babynames, pipeline2)
```



The transformations can also be initialized as base classes, allowing for them to be shared between pipelines:

``` {r, eval=F}
psu <- Pseudonymizer$new()

pipeline3 <- deident(psu, Var1, Var2)
pipeline4 <- deident(psu, Var3, Var4)

```


Performing a multi-stage transformation can be done by chaining together calls to `deident`

``` {r, eval=F}
blur <- NumericBlurrer$new(cuts=c(0, 10, 100, 1000, 10000))

pipeline3 <- deident(babynames, "psudonymize", name, sex) |>
  deident(blur, n)

apply_deident(babynames, pipeline3)
```

An in depth example can be found in the `Worked Example` vigentte, while 
more details on each method are presented in the `deident transforms` vignette.

## Current applications

The `deident` toolbox was developed for applications in the NuRS and AmreS research projects which aim to extract and analyze retrospective operational data from NHS Trusts to understand staff retention and patient safety.  

## Contributions

The package was designed by RC, MA, and SJ.  Implementation was done by RC. Quality assurance was done by MA.  Documentation was written by RC, and SJ.  Funding for the work was won by RC and SJ.

## Acknowledgements

The development of `deident` was part of the he NuRS and AmReS projects funded by the Health Foundation.

## References

