---
output: 
  md_document:
    variant: markdown_github
---

Research in the discipline of health data is of increasing interest due to the 
perception that artificial intelligence (AI) and machine learning (ML) 
techniques have the promise to bridge gaps in the large, resource-limited 
sector.  One key aspect of working within health data is the necessity to work 
within the ethical and legal frameworks of research with human participants, 
notably around the risks posed by the processing of personally identifiable 
data (PID) and pseudo-PID.  In the UK, there has been a drive to increase 
adoption of open source software into the working practices of the NHS, 
corresponding to the growth on the NHS-R community and similar groups.  
Here we demonstrate an extendable package of tools for the implementation 
and application of deidentification techniques to panel datasets.

This package implements methods for de-identification via:

*	pseudonymization – the consistent replacement of a string by a random string
*	encryption 
*	shuffling – replacement of columns by a random sample without replacement
*	bluring – the aggregation of numeric or categorical data according to specified rules
*	perturbation – the addition of user-defined random noise to a numeric variable

following the design principles of the existing tidyverse domain for ease of 
adoption.  The package includes tools to create a single pipeline for the 
application of a multi-step deidentification pipeline to multiple files 
stored within the same repository, alongside the option to serialize/ define 
the pipeline to/ via yaml.  This approach allows a researcher to design and 
implement an appropriate de-identification plan and deliver it to the research 
support/ business intelligence team of an organisation with limited knowledge 
of the sensitive data.  The supply of an easy method for de-identifying data
sets which require little scripting knowledge by Trust staff may aid in 
overcoming several information governance risks that keep operational data 
siloed within health Trusts.  

The core functionality of the package is the `deident` function. To demonstrate functionality we use a subset of the `babynames` data set consisting of the final two years.  

``` {r}
library(deident)

babynames <- babynames::babynames |> 
  dplyr::filter(year > 2015) |> 
  data.frame()

str(babynames)
```

The `deident` function produces a pipeline of actions and the variables to be transformed, with each subsequent call of `deident` adding a further transformation.    The first argument of `deident` can either be a data frame or the output of a previous `deident` call.

The simplest use case is  transforming one variable via a single method:

``` {r, eval=F}
pipeline <- deident(babynames, "psudonymize", name)

apply_deident.data.frame(babynames, pipeline)

```

The same method can be applied to multiple variables by adding the variable names:

``` {r, eval=F}
pipeline2 <- deident(babynames, "psudonymize", name, sex)

apply_deident.data.frame(babynames, pipeline2)
```



The transformations can also be initialized as base classes, allowing for them to be shared between pipelines:

``` {r, eval=F}
psu <- Pseudonymizer$new()

pipeline3 <- deident(psu, Var1, Var2)
pipeline4 <- deident(psu, Var3, Var4)

```


Performing a multi-stage transformation can be done by chaining together calls to `deident`

``` {r, eval=F}
blur <- NumericBlurer$new(cuts=c(0, 10, 100, 1000, 10000))

pipeline3 <- deident(babynames, "psudonymize", name, sex) |>
  deident(blur, n)

apply_deident.data.frame(babynames, pipeline3)
```

For more details on each method refer to the `deident transforms` vignette.




