---
title: "Perturb Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Perturb Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Not all sensitive data is recorded as strings - features such as age, date of birth, or income could result in aspects of a data set being personally identifiable.  To aid with these challenges we include methods for 'perturbing' numeric data (the addition of random noise). 

Three types of random noise are included:

1. `adaptive_noise` [default] - random noise which scales with the standard deviation of the variable transformed.
2. `white_noise` - random noise at a set spread.
3. `lognorm_noise` - random multiplicative noise at a set spread.

NB: we set a random seed using `set.seed` here for reproducibility.  We recommend users avoid this step when using the package in production code.

```{r setup}
library(deident)
set.seed(101)

perturb_pipe <- ShiftsWorked |>
  add_perturb(`Daily Pay`)

apply_deident(ShiftsWorked, perturb_pipe)
```

To change the noise, pass one of the functions including the desired level of noise.

```{r}
perturb_pipe_white_noise <- ShiftsWorked |>
  add_perturb(`Daily Pay`, noise = white_noise(sd = 0.3))

apply_deident(ShiftsWorked, perturb_pipe_white_noise)
```

```{r }
perturb_pipe_heavy_adaptive_noise <- ShiftsWorked |>
  add_perturb(`Daily Pay`, noise = adaptive_noise(sd.ratio = 0.4))

apply_deident(ShiftsWorked, perturb_pipe_heavy_adaptive_noise)
```
