---
title: "Re-using Methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reusing_methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

NB: the following is an advanced usage of `deident`.  If you are just getting started we recommend looking at the other vignettes first.

While the `deident` package implements multiple different methods for deidentification, one of its key advantages is the ability to re-use and share methods across data sets due to the 'stateful' nature of its design.

If you wish to share a unit between different pipelines, the cleanest approach is to initialize the method of interest and then pass it into the first pipeline:

```{r setup}
library(deident)

psu <- Pseudonymizer$new()

name_pipe <- starwars |>
  deident(psu, name)

apply_deident(starwars, name_pipe)
```

Having called `apply_deident` the Pseudonymizer `psu` has learned encodings for each string in `starwars$name`.  If these strings appear a second time, they will be replaced in the same way, and we can build a second pipeline using `psu`: 

``` {r}
combined.frm <- data.frame(
  ID = c(head(starwars$name, 5), head(ShiftsWorked$Employee, 5))
)

reused_pipe <- combined.frm |>
  deident(psu, ID)

apply_deident(combined.frm, reused_pipe)

```

Since the first 5 lines of `combined.frm$ID` are the same as `starwars$ID` the first 5 lines of each transformed data set are also the same.

