---
title: "Transformations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transformations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Out of the box, `deident` features a set of transformations to aid in the de-identification of data sets.   Each transformation is implemented via `R6Class`  and extends `BaseDeident`.  User defined  transformations can be implemented in a similar manner.

To demonstrate the different transformation we supply a toy data set, `df`, comprising 26 observations of three variables:

* A:  character, a to z
* B: numeric, 1 to 26
* C: character, `X` if `B <= 13`, `Y` if `B > 13`

``` {r, include=F}
library(deident)

df <- data.frame(
  A = letters,
  B = 1:26,
  C = sort(rep(c("X", "Y"), 13))
)
df
```

## Psudonymizer

Apply a cached random replacement cipher.  Re-occurrence of the same key will receive the same hash.

Implemented `deident` options:

``` {r, eval=F}
deident(df, "psudonymize", A)
deident(df, "Pseudonymizer", A)
deident(df, Pseudonymizer, A)
deident(df, Pseudonymizer$new(), A)

psu <- Pseudonymizer$new()
deident(df, psu, A)
```

### Options

By default `Pseudonymizer` replaces  values in variables with a random alpha-numeric string of 5 characters.  This can be replaced via calling `set_method` on an instantiated Pseudonymizer with the desired function:

``` {r}
psu <- Pseudonymizer$new()

new_method <- function(key, ...) {
  paste(sample(letters, 12, T), collapse = "")
}

psu$set_method(new_method)

deident(df, psu, A)
```

The first argument to the method receives the key to be transformed.

## Shuffler

Implemented `deident` options:

``` {r, eval=F}
deident(df, "shuffle", A)
deident(df, "Shuffler", A)
deident(df, Shuffler, A)
deident(df, Shuffler$new(), A)

shuffle <- Shuffler$new()
deident(df, shuffle, A)
```

## Encrypter

Apply cryptographic hashing to a variable.

Implemented `deident` options:

``` {r, eval=F}
deident(df, "encrypt", A)
deident(df, "Encrypter", A)
deident(df, Encrypter, A)
deident(df, Encrypter$new(), A)

encrypt <- Encrypter$new()
deident(df, encrypt, A)
```

### Options

At initialization, `Encrypter` can be given `hash_key` and `seed` values to control the cryptographic encryption.  It is recommended users set these values and do not disclose them.

``` {r}
encrypt <- Encrypter$new(hash_key = "deident_hash_key_123", seed = 202)
deident(df, encrypt, A)
```

## Perturber

Apply Gaussian white noise to a numeric variable.

Implemented `deident` options:

``` {r, eval=F}
deident(df, "perturb", A)
deident(df, "Perturber", A)
deident(df, Perturber, A)
deident(df, Perturber$new(), A)

perturb <- Perturber$new()
deident(df, perturb, A)
```

### Options

At initialization, `Perturber` can be given a scale for the white noise via the `sd` argument.  

``` {r}
perturb <- Perturber$new(noise = adaptive_noise(0.2))
deident(df, perturb, B)
```

## Blurrer

Aggregate categorical values dependent on a user supplied list.  the list must be supplied to `Blur` at initialization.

Implemented `deident` options:

``` {r, eval=F}
letter_blur <- c(rep("Early", 13), rep("Late", 13))
names(letter_blur) <- letters

blur <- Blurrer$new(blur = letter_blur)
deident(df, blur, A)
```

## NumericBlurrer

Aggregate numeric values dependent on a user supplied vector of breaks/ cuts.  If no vector is supplied `NumericBlurrer` defaults to a binary classification about 0.

Implemented `deident` options:

``` {r, eval=F}
deident(df, "numeric_blur", B)
deident(df, "NumericBlurrer", B)
deident(df, NumericBlurrer, B)
deident(df, NumericBlurrer$new(), B)

numeric_blur <- NumericBlurrer$new()
deident(df, numeric_blur, B)
```
### Options

At initialization  `NumericBlurrer` takes an argument `cuts` to define the limits of each interval.

``` {r}
numeric_blur <- NumericBlurrer$new(cuts = c(5, 10, 15, 20))
deident(df, numeric_blur, B)
```

## GroupedShuffler

Apply `Shuffler` to a data set having first grouped the data on column(s).  The grouping needs to be defined at initialization.

Implemented `deident` options:

``` {r, eval=F}
grouped_shuffle <- GroupedShuffler$new(C)
deident(df, grouped_shuffle, B)
```
### Options

At initialization  `GroupedShuffler` takes an argument `limit` such that if any aggregated sub group has fewer than `limit` observations all values are dropped.

``` {r}
numeric_blur <- GroupedShuffler$new(C, limit = 1)
deident(df, numeric_blur, B)
```

## Drop

Define a column to be removed from the pipeline.

Implemented `deident` options:

``` {r, eval=F}

deident(df, Drop, B)

drop <- deident:::Drop$new()
deident(df, drop, B)
```
