---
title: "Transformations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transformations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.print_min = 4L, tibble.print_max = 4L)
```

Out of the box, `deident` features a set of transformations to aid in the de-identification of data sets.   Each transformation is implemented via `R6Class`  and extends `BaseDeident`.  User defined  transformations can be implemented in a similar manner.

To demonstrate the different transformation we use the `babynames` data set co-opted from the `babynames` package:  

``` {r} 
library(deident)

babynames <- babynames::babynames |> 
  dplyr::filter(year > 2015) 

babynames
```
## Psudonymizer

Apply a cached random replacement cipher.  Re-occurrence of the same key will receive the same hash.

Implemented `deident` options:

``` {r, eval=F}
# Examples only:
deident(df, "psudonymize", ...)
deident(df, "Pseudonymizer", ...)
deident(df, Pseudonymizer, ...)
deident(df, Pseudonymizer$new(), ...)

psu <- Pseudonymizer$new()
deident(df, psu, ...)
```

### Options

By default `Pseudonymizer` replaces  values in variables with a random alpha-numeric string of 5 characters.  This can be replaced via calling `set_method` on an instantiated Pseudonymizer with the desired function:

``` {r, result="asis"}
psu <- Pseudonymizer$new()

new_method <- function(key, ...) {
  paste(sample(letters, 12, T), collapse = "")
}

psu$set_method(new_method)

dlist_psu <- deident(babynames, psu, name)
dlist_psu

apply_deident(babynames, dlist_psu)
```

The first argument to the method receives the key to be transformed.

## Shuffler

Implemented `deident` options:

``` {r, eval=F}
# Examples only:
deident(df, "shuffle", ...)
deident(df, "Shuffler", ...)
deident(df, Shuffler, ...)
deident(df, Shuffler$new(), ...)

shuffle <- Shuffler$new()
deident(df, shuffle, ...)
```

## Encrypter

Apply cryptographic hashing to a variable.

Implemented `deident` options:

``` {r, eval=F}
# Examples only:
deident(df, "encrypt", ...)
deident(df, "Encrypter", ...)
deident(df, Encrypter, ...)
deident(df, Encrypter$new(), ...)

encrypt <- Encrypter$new()
deident(df, encrypt, ...)
```

### Options

At initialization, `Encrypter` can be given `hash_key` and `seed` values to control the cryptographic encryption.  It is recommended users set these values and do not disclose them.

``` {r}
encrypt <- Encrypter$new(hash_key = "deident_hash_key_123", seed = 202)
dlist_enc <- deident(babynames, encrypt, name)
dlist_enc

apply_deident(babynames, dlist_enc)
```

## Perturber

Apply Gaussian white noise to a numeric variable.

Implemented `deident` options:

``` {r, eval=F}
# Example only:
deident(df, "perturb", ...)
deident(df, "Perturber", ...)
deident(df, Perturber, ...)
deident(df, Perturber$new(), ...)

perturb <- Perturber$new()
deident(df, perturb, ...)
```

### Options

At initialization, `Perturber` can be given a scale for the white noise via the `sd` argument.  
``` {r}
perturb <- Perturber$new(noise = adaptive_noise(0.2))
dlist_pert <- deident(babynames, perturb, prop)
dlist_pert

apply_deident(babynames, dlist_pert)
```

## Blurrer

Aggregate categorical values dependent on a user supplied list.  the list must be supplied to `Blur` at initialization.

Implemented `deident` options:

``` {r, eval=F}
# Example only:
letter_blur <- c(rep("Early", 13), rep("Late", 13))
names(letter_blur) <- letters

blur <- Blurrer$new(blur = letter_blur)
deident(df, blur, A)
```

## NumericBlurrer

Aggregate numeric values dependent on a user supplied vector of breaks/ cuts.  If no vector is supplied `NumericBlurrer` defaults to a binary classification about 0.

Implemented `deident` options:

``` {r, eval=F}
# Example only:
deident(df, "numeric_blur", ...)
deident(df, "NumericBlurrer", ...)
deident(df, NumericBlurrer, ...)
deident(df, NumericBlurrer$new(), ...)

numeric_blur <- NumericBlurrer$new()
deident(df, numeric_blur, ...)
```
### Options

At initialization  `NumericBlurrer` takes an argument `cuts` to define the limits of each interval.

``` {r}
numeric_blur <- NumericBlurrer$new(cuts = c(10, 30))
dlist_nb <- deident(babynames, numeric_blur, n)
dlist_nb

apply_deident(babynames, dlist_nb)
```

## GroupedShuffler

Apply `Shuffler` to a data set having first grouped the data on column(s).  The grouping needs to be defined at initialization.

Implemented `deident` options:

``` {r, eval=F}
# Example only:
grouped_shuffle <- GroupedShuffler$new(year)
deident(babynames, grouped_shuffle, name)
```
### Options

At initialization  `GroupedShuffler` takes an argument `limit` such that if any aggregated sub group has fewer than `limit` observations all values are dropped.

``` {r}
numeric_blur <- GroupedShuffler$new(year, limit = 1)
dlist_groupshuffle <- deident(babynames, numeric_blur, name)
dlist_groupshuffle

apply_deident(babynames, dlist_groupshuffle)
```

## Drop

Define a column to be removed from the pipeline.

Implemented `deident` options:

``` {r, eval=F}
# Example only:
deident(df, Drop, ...)

drop <- deident:::Drop$new()
deident(df, drop, ...)
```
